\Large\textbf{Preparation GUIDE for Initial-Submission Manuscript}\\~\\ \large\textbf{ Preparation of Papers in Two-Column Format\\ for the ASP-DAC 2014 (\LaTeX2e version)}

### 2014-01-27 conclusion.tex
Added proofs, can talk a bit more about completeness and soundness

In this paper we have described a novel way to introduce real-time in
\textit{Globally Asynchronous Locally Synchronous} (GALS) languages, in
particular SystemJ, and their subset the synchronous languages like
Esterel. The fundamental idea is to convert real-time wait constructs
into logical waits (\texttt{pause} construct) that interact well with
rest of the constructs in these languages, especially, preemption and
channel communication. We do \textit{not} need or use external timers to
introduce these real-time wait constructs, instead we depend upon an
underlying platform that is timing analyzable, thereby resolving the
problems of interaction between external timers and
preemption. Moreover, we also do not require specific timer resolutions,
thereby making our solution elegant (in the spirit of GALS/synchronous
programming) and flexible. We rewrite the wait constructs into reactive
kernel constructs, calculate the \textit{Worst} and \textit{Best} case
reaction times and then determine wait \texttt{d} in number of logical
ticks, which is invariant to the \textit{elasticity} of logical time. To
our knowledge we are the first to introduce exact and inexact wait
constructs in such languages and show their usefulness in designing
real-world systems.

>>>>

In this paper we have described a novel way to introduce real-time in
\textit{Globally Asynchronous Locally Synchronous} (GALS) languages, in
particular SystemJ, and their subset the synchronous languages like
Esterel. The fundamental idea is to convert real-time wait constructs
into logical waits (\texttt{pause} construct) that interact well with
rest of the constructs in these languages, especially, preemption and
channel communication. We do \textit{not} need or use external timers to
introduce these real-time wait constructs, instead we depend upon an
underlying platform that is timing analyzable, thereby resolving the
problems of interaction between external timers and
preemption. Moreover, we also do not require specific timer resolutions,
thereby making our solution elegant (in the spirit of GALS/synchronous
programming) and flexible. We rewrite the wait constructs into reactive
kernel constructs, calculate the \textit{Worst} and \textit{Best} case
reaction times and then determine wait \texttt{d} in number of logical
ticks, which is invariant to the \textit{elasticity} of logical time. To
our knowledge we are the first to introduce exact and non-exact wait
constructs in such languages and show their usefulness in designing
real-world systems.

### 2014-01-27 delays.tex
Added proofs, can talk a bit more about completeness and soundness

We also introduce two variants of the derived statement
\texttt{wait\_inbetween}.
\begin{enumerate*}
\item Given a SystemJ program \texttt{wait\_atleast (M); p}, where $M
  \in \mathbb{Q}^{>0}$, statement $p$ is executed after real-time delay
  of $\tau$ units, such that, $M \leq \tau < \infty$. It is important to
  note that the lower bound of the wait construct; $M$ is \textit{not}
  an exact delay, but rather the control is allowed to proceed to the
  next statement anytime after waiting for \textit{at least} time $M$.
\item Given a SystemJ program \texttt{wait\_exactly (M); p}, where $M
  \in \mathbb{Q}^{>0}$, statement $p$ is executed after real-time delay
  of $\tau$ units, such that, $M \leq \tau \leq M$. In this variant, the
  waiting time is exact.
\end{enumerate*}

@@

Let us revisit our motivating example to illustrate the algorithm. From
Figure~\ref{fig:1} we know that $M$ is 50.3 ms and $N$ is 200.3 ms,
respectively. Let the WCRT and BCRT be: 0.112 ms and 0.0334 ms,
respectively. Thus, the algorithm proceeds as follows:

\begin{enumerate*}
\item $l_1 \leftarrow \lceil 50.3/0.112 \rceil$ and $u_1 \leftarrow
  \lfloor 200.3/0.112 \rfloor$. $l_1 = 450$ and $u_1 = 1788$. We first
  calculate the logical ticks that are always running at the WCRT and
  satisfy the required real-time wait.
\item $l_2 \leftarrow \lceil 50.3/0.0334 \rceil$ and $u_2 \leftarrow
  \lfloor 200.3/0.0334 \rfloor$. $l_1 = 1506$ and $u_1 = 5997$. We do
  the same for the BCRT case.
\item $S_1 = [450,1788]$ and $S_2 =[1506,5997]$. We then map the
  resultant bounds to linear points. Sets $S_1$ and $S_2$ represent
  logical ticks that running at the WCRT and BCRT, respectively always
  satisfy the required real-time constraints.
\item $D = S_1 \cap S_2$, $D = [1506,1788]$. Finally, the intersection
  of the two sets gives the set $D$ from which we can choose any value
  for \texttt{d}.
\end{enumerate*}

@@

The resultant value for \texttt{d} gives the number of logical ticks,
which can run at any physical clock-speed, bounded by the BCRT and the
WCRT of the program and still result in the desired real-time
wait. Figure~\ref{fig:hrtcssoln} gives the solution space for the
aforementioned calculation. The patterned space is the only overlapping
region and hence gives the solution space for the above example.

@@

\paragraph{Period execution of the clock domains}
\label{sec:peri-exec-clock}

The aforementioned technique maintains the essence of synchronous and
GALS programs -- the logical tick, but in doing so is unable to provide
the exact real-time wait of $M$ time units in case of
\texttt{wait\_exactly(M)} statement, thereby violating its
semantics. The second option is to execute the CD periodically at the
WCRT. WCRT determines the execution speed of any CD guaranteeing that no
input events from the environment will be missed
(cf. Section~\ref{sec:mapping-logical-time}). Thus, if we simply execute
the CD at periodically at the WCRT, we can choose the value of
\textit{d} from set $S1$. This technique allows meeting the synchronous
hypothesis, while at the same time guaranteeing the semantics of the
\texttt{wait\_exactly(M)} statement. But, compared to the previous
option, this technique is very inflexible and results in reduced
response time and increased power consumption.

We believe that the designer should be allowed to choose either of these
techniques and hence, provide this option via compiler switches.

@@

Our compiler has a switch that enables over-approximation as described
in Section~\ref{sec:extend-tehcn-vari} provided the wait requirements
are not met. If this compiler switch is not enabled, the compiler simply
decides to execute the CD at the periodically, where the period is the
WCRT. The system designer can then either set the compiler switch or
change the program by shortening the critical path by introducing
\texttt{pause} constructs. We show how introducing \texttt{pause}
constructs can help reduce the WCRT in
Section~\ref{sec:experimental-results}.

>>>>

We also introduce two variants of the derived statement
\texttt{wait\_inbetween}.
\begin{enumerate*}
\item Given a SystemJ program \texttt{wait\_atleast (M); p}, where $M
  \in \mathbb{Q}^{>0}$, statement $p$ is executed after real-time delay
  of $\tau$ units, such that, $M \leq \tau < \infty$. It is important to
  note that the lower bound of the wait construct; $M$ is \textit{not}
  an exact delay, but rather the control is allowed to proceed to the
  next statement anytime after waiting for \textit{at least} $M$ times
  units.
\item Given a SystemJ program \texttt{wait\_exactly (M); p}, where $M
  \in \mathbb{Q}^{>0}$, statement $p$ is executed after real-time delay
  of $\tau$ units, such that, $M \leq \tau \leq M$. In this variant, the
  waiting time is exact.
\end{enumerate*}

@@

Let us revisit the HRTCS example (cf. Figure~\ref{fig:1}) to illustrate
the algorithm. From Figure~\ref{fig:1} we know that $M$ is 50.3 ms and
$N$ is 200.3 ms, respectively. Let the WCRT and BCRT be: 0.112 ms and
0.0334 ms, respectively. Thus, the algorithm proceeds as follows:

\begin{enumerate*}
\item $l_1 \leftarrow \lceil 50.3/0.112 \rceil$ and $u_1 \leftarrow
  \lfloor 200.3/0.112 \rfloor$. $l_1 = 450$ and $u_1 = 1788$. We first
  calculate the logical ticks that are always running at the WCRT and
  satisfy the required real-time wait period.
\item $l_2 \leftarrow \lceil 50.3/0.0334 \rceil$ and $u_2 \leftarrow
  \lfloor 200.3/0.0334 \rfloor$. $l_2 = 1506$ and $u_2 = 5997$. We do
  the same for the BCRT case.
\item $S_1 = [450,1788]$ and $S_2 =[1506,5997]$. We then map the
  resultant bounds to linear points. Sets $S_1$ and $S_2$ represent
  logical ticks that running at the WCRT and BCRT, respectively always
  satisfy the required real-time constraints.
\item $D = S_1 \cap S_2$, $D = [1506,1788]$. Finally, the intersection
  of the two sets gives the set $D$ from which we can choose any value
  for \texttt{d}.
\end{enumerate*}

@@

The resultant value for \texttt{d} gives the number of logical ticks,
which can run at any physical clock-speed, bounded by the BCRT and the
WCRT of the program and still result in the desired real-time wait
period. Figure~\ref{fig:hrtcssoln} gives the solution space for the
aforementioned calculation. The patterned space is the only overlapping
region and hence gives the solution space for the above example.

@@

\paragraph{Periodic execution of the clock domains}
\label{sec:peri-exec-clock}

The aforementioned technique maintains the essence of synchronous and
GALS programs -- the logical tick, but in doing so is unable to provide
the exact real-time wait of $M$ time units in case of
\texttt{wait\_exactly(M)} statement, thereby violating its
semantics. The second option is to execute the CD periodically at the
WCRT. WCRT determines the fastest execution speed of any CD with the
guarantee that no input events from the environment will be missed
(cf. Section~\ref{sec:mapping-logical-time}). Thus, if we simply execute
the CD at periodically at the WCRT, we can choose the value of
\textit{d} from set $S1$. This technique allows meeting the synchronous
hypothesis, while at the same time guaranteeing the semantics of the
\texttt{wait\_exactly(M)} statement. But, compared to the previous
option, this technique is very inflexible and results in reduced
response time and increased power consumption.

We believe that the designer should be allowed to choose from either of
these techniques and hence, provide this option via compiler switches.

@@

Our compiler has a switch that enables over-approximation as described
in Section~\ref{sec:extend-tehcn-vari} provided the wait requirements
are not met. If this compiler switch is not enabled, the compiler simply
decides to execute the CD periodically, where the period is the
WCRT. WCRT effects the over-approximation or the period of the CD,
irrespective of the solution that the system designer chooses. A large
WCRT results in reduced performance and hence it is essential that the
WCRT be optimized. One can optimize the WCRT of any CD by breaking the
critical paths using the \texttt{pause} construct as we show later in
Section~\ref{sec:experimental-results}.

% The system designer can then either set the compiler switch or change
% the program by shortening the critical path by introducing
% \texttt{pause} constructs. We show how introducing \texttt{pause}
% constructs can help reduce the WCRT in
% Section~\ref{sec:experimental-results}.

### 2014-01-27 discussion.tex
Added proofs, can talk a bit more about completeness and soundness

The system designer needs to be aware that just introducing a wait
construct does not make the reaction (or CD) periodic. The program on
the left does \textit{not} emit signal \texttt{S} every \texttt{1 ms},
because of the extra \texttt{pause} statement. Introducing additional
pauses introduces more logical ticks, consequently increasing the
real-time wait constructs. Static analysis techniques from real-time
community can be used to automatically determine the wait specification,
in such cases, but describing these techniques is out of the scope of
this work.

@@

As mentioned previously, the proposed solution might need
over-approximation techniques (Algorithm~\ref{alg:2}) to guarantee
timing requirements (cf. Section~\ref{sec:extend-tehcn-vari}). This
might give the readers an impression that timer based solution proposed
for standard real-time operating systems should be utilized for
GALS/synchronous programs. This is a fallacy. Bourke et
al~\cite{Bourke2009a} have investigated this technique
unsuccessfully. We elaborate upon the interplay of timers and reactivity
to make the problem of using timers explicit.

Preemption plays an important role in reactive languages. One needs to
carefully consider the interplay of wait constructs semantics with the
preemption semantics of reactive languages. Previous attempts at
incorporating delay (using external timers) have only been partially
successful, because of the complex interplay between real-time and
preemption. Consider the simple example below, which models real-time
using external timers as in~\cite{rsh94}.

@@

As identified in~\cite{Bourke2009a} \texttt{suspend} does not play well
with the external timer. The above program sends a signal to an external
timer and waits for 10 ms to pass by. Consider what happens when signals
\texttt{S} and \texttt{TIMER} occur in the same logical tick, the
\texttt{await} statement is never executed (due to suspend) and hence,
we enter a deadlock. Such problems are completely avoided in our
technique, because we convert the real-time wait constructs into logical
waits (\texttt{pause} constructs), which interact well with preemption.
Furthermore, synchronous/GALS languages have always been targeted at
real-time analyzable platforms and constrained
environments~\cite{DBLP:journals/pieee/SifakisTY03,boldt07}. One can use
our technique within such a setting with ease.

@@

\begin{figure}[h!]
  \centering
	\vspace{-10pt}
		\begin{lstlisting}[style=sysj,basicstyle=\normalsize\ttfamily,morekeywords={emit,trap,pause,exit,delay,suspend}]
{
 //reaction R1
 emit START_TIMER(1 ms); 
 await (TIMER);emit O1;
}
||
{
 //reaction R2
 // do something
}
		\end{lstlisting}
  \caption{Resolution of external timers}
  \label{fig:resolution}
	\vspace{-10pt}
\end{figure}

There is yet another problem with timer based systems. Let us consider
the program in Figure~\ref{fig:resolution}. On a cursory look this
example should work fine. An external timer is started in reaction R1,
which counts down from 1 ms. Once this time has elapsed a signal
\texttt{TIMER} is generated from this external timer, which should emit
O1 in turn. There is no \texttt{suspend} construct encapsulating the
timer and hence all seems fine. Let us now consider what happens to the
whole CD, including reaction R2; while reaction R1 is waiting for the
\texttt{TIMER} signal, reaction R2 is making progress. Input signals can
only be captured from the environment at the logical tick boundaries
(cf. Section~\ref{sec:mapping-logical-time}). In
Figure~\ref{fig:resolution}, reaction R2 determines the length of the
tick, because it is performing the heavier computation. Consider what
happens if the external timer generates the \texttt{TIMER} signal, but
the system is still performing computation and the logical tick has not
yet finished -- we \textit{miss} this \texttt{TIMER} signal from the
environment. Thus, even in the absence of the \texttt{suspend}
construct, we are not guaranteed that the program will capture the
external timer signal. In fact, the WCRT needs to be smaller than 1 ms
for the above program to perform as expected. But, this again implies
need for a time analyzable platform even when using external timers.

In general we observe that the WCRT determines the lowest resolution of
any real-time construct whether it be external timer independent, like
the proposed technique, or external timer dependent, like above. This
discovery is one of the key insights for proposing the solution
described in this paper.

>>>>

The system designer needs to be aware that just introducing a wait
construct does not make the reaction (or CD) periodic. The program on
the left does \textit{not} emit signal \texttt{S} every \texttt{1 ms},
because of the extra \texttt{pause} statement. Introducing additional
pauses introduces more logical ticks, consequently increasing the
real-time wait period. Static analysis techniques from real-time
community can be used to automatically determine the wait specification,
in such cases, but describing these techniques is out of the scope of
this work.

@@

% As mentioned previously, the proposed solution might need
% over-approximation techniques (Algorithm~\ref{alg:2}) to guarantee
% timing requirements (cf. Section~\ref{sec:extend-tehcn-vari}). This
% might give the readers an impression that timer based solution proposed
% for standard real-time operating systems should be utilized for
% GALS/synchronous programs. This is a fallacy. Bourke et
% al~\cite{Bourke2009a} have investigated this technique
% unsuccessfully. We elaborate upon the interplay of timers and reactivity
% to make the problem of using timers explicit.

Preemption plays an important role in reactive languages. One needs to
carefully consider the interplay of wait constructs semantics with the
preemption semantics of reactive languages. Previous attempts at
incorporating postponement using external timers has only been partially
successful, because of the complex interplay between real-time and
preemption. Consider the simple example below, which models real-time
using external timers as in~\cite{rsh94}.

@@

As identified in~\cite{Bourke2009a} \texttt{suspend} does not play well
with the external timer. The program in Figure~\ref{fig:preemp} sends a
signal to an external timer and waits for 10 ms to pass by. Consider
what happens when signals \texttt{S} and \texttt{TIMER} occur in the
same logical tick, the \texttt{await} statement is never executed (due
to suspend) and hence, we enter a deadlock. Such problems are completely
avoided in our technique, because we convert the real-time wait
constructs into logical waits (\texttt{pause} constructs), which
interact well with preemption.  Furthermore, synchronous/GALS languages
have always been targeted at real-time analyzable platforms and
constrained
environments~\cite{DBLP:journals/pieee/SifakisTY03,boldt07}. One can use
our technique within such a setting with ease.

@@

\begin{figure}[h!]
  \centering
	\vspace{-10pt}
		\begin{lstlisting}[style=sysj,basicstyle=\normalsize\ttfamily,morekeywords={emit,trap,pause,exit,delay,suspend}]
{
 //reaction R1
 emit START_TIMER(1 ms); 
 await (TIMER);emit O1;
}
|| //synchronous parallel
{
 //reaction R2
 // do something
}
		\end{lstlisting}
  \caption{Resolution of external timers}
  \label{fig:resolution}
	\vspace{-10pt}
\end{figure}

There is yet another problem with timer based systems. Let us consider
the program in Figure~\ref{fig:resolution}. On a cursory look this
example should work fine. An external timer is started in reaction R1,
which counts down from 1 ms. Once this time has elapsed, a signal
\texttt{TIMER} is generated from this external timer, which should emit
O1 in turn. There is no \texttt{suspend} construct encapsulating the
timer and hence all seems fine. Let us now consider what happens to the
whole CD, including reaction R2; while reaction R1 is waiting for the
\texttt{TIMER} signal, reaction R2 is making progress. Input signals can
only be captured from the environment at the logical tick boundaries
(cf. Section~\ref{sec:mapping-logical-time}). In
Figure~\ref{fig:resolution}, reaction R2 determines the length of the
tick, because it is performing the heavier computation. Consider what
happens if the external timer generates the \texttt{TIMER} signal, but
the system is still performing computation and the logical tick has not
yet finished -- we \textit{miss} this \texttt{TIMER} signal from the
environment. Thus, even in the absence of the \texttt{suspend}
construct, we are not guaranteed that the program will capture the
external timer signal. In fact, the WCRT needs to be smaller than 1 ms
for the above program to perform as expected. But, this again implies
need for a time analyzable platform even when using external timers.

In general we observe that the WCRT determines the lowest resolution of
any real-time construct whether it be external timer independent, like
the proposed technique, or external timer dependent, like above. This
discovery is one of the key insights for proposing the solutions
described in this paper.

### 2014-01-27 exps.tex
Added proofs, can talk a bit more about completeness and soundness

We have carried out a number of experiments on a set of applications
with real-time constraints. The characteristics of the benchmark set
including required real-time delays and program sizes are shown in
Table~\ref{fig:benchlist}.
{\renewcommand{\arraystretch}{0.9}
%\renewcommand{\texttt}[1] {\begin{lstlisting}[language=Java,morekeywords={delay}] #1 \end{lstlisting}}
\begin{table}[h!]
	\scriptsize
	\centering
	\caption{An overview of the benchmark programs}
	\label{fig:benchlist}
	\scalebox{0.9}{
		\begin{tabular}{llp{0.5cm}p{1cm}}
\toprule
\emph{Benchmark}       & \emph{Delay constructs used in each CD} & \emph{\# of CDs} & 
\emph{Lines of code}        
\\ \midrule
HRTCS                  & \texttt{wait\_inbetween(50.3..200.3 ms)} ($CD1$) & 2                  & 56
\\ \midrule
\multirow{5}{*}{Motor} & \texttt{wait\_exactly(2.4 ms)}        & \multirow{5}{*}{1} & \multirow{5}{*}{116} \\ 
											 &
                                                                                         \texttt{wait\_exactly(1.667
                                                                                           ms)}      &                    &                      \\ 
											 &
                                                                                    \texttt{wait\_exactly(0.05
                                                                                           ms)}       &                    &                      \\ 
											 &
                                                                                    \texttt{wait\_exactly(0.3
                                                                                           ms)}        &                    &                      \\ 
											 &
                                                                                         \texttt{wait\_exactly(0.733
                                                                                           ms)}      &                    &
\\ \midrule
\multirow{3}{*}{Printer}
& \texttt{wait\_inbetween(2.7..10.6 ms)} ($CD1$)      & \multirow{3}{*}{2}        & \multirow{3}{*}{39} \\
& \texttt{wait\_inbetween(1.2..30.3 ms)} ($CD2$)		  &														&											\\
& \texttt{wait\_inbetween(5.6..100.2 ms)} ($CD2$)     &														&											
\\ \midrule
\multirow{2}{*}{AECS} & \texttt{wait\_atleast(10000 ms)} ($CD1$) 
& \multirow{2}{*}{2} & \multirow{2}{*}{267} \\
& \texttt{wait\_exactly(10000 ms)} ($CD2$) & &
\\ \bottomrule
\end{tabular}
}
\end{table}
}
{\renewcommand{\arraystretch}{0.6}
\begin{table*}[t!]
\centering
	\caption{Actual delays obtained for wait constructs in the
benchmark programs based on their BCRT and WCRT}
\begin{tabular}{ c c l l l l l l }
	\toprule
	\multicolumn{2}{c}{} & \emph{HRTCS} & \emph{Motor} & \emph{Printer/CD1} &
	\emph{Printer/CD2} & \emph{AECS/CD1} &	\emph{AECS/CD2} \\ 
	\midrule
	\multicolumn{2}{c}{\emph{BCRT}}	& 0.0.0396 ms	& 0.0425 ms & 0.0190 ms &
	0.0263 ms & 0.1784 ms & 0.1284 ms\\ 
	\cmidrule(r){1-8}
	\multicolumn{2}{c}{\emph{WCRT}} &0.1152 ms & 0.1096 ms & 0.0647 ms
	& 0.0491 ms & 0.3476 ms & 0.2991 ms\\ 
	\cmidrule(r){1-8}

@@

AECS (Access and Environmental Control System) is an application that
controls an intelligent room environment described
in~\cite{aecs_ispa}. The system interacts with two external timers
through input/output signals in order to measure how long the entrance
door to the room is being opened to detect intruders.  The program has
been modified such that the wait constructs are used instead of external
timers as shown in Figure~\ref{delaycode}. In the original AECS, the
system needed to trigger external timer by emitting the output signal in
order to wait for a specific amount of time (e.g.
\texttt{TimerTriggered} in Figure~\ref{delaycode:a}). When the timer
expires, it notifies the system through the input signal
\texttt{TimeOut}, which is polled using the \texttt{await}
statement. With the new wait constructs, these timers are not necessary
and waits can be expressed directly in the code as shown in
Figure~\ref{delaycode:b}.

>>>>

We have carried out a number of experiments on a set of applications
with real-time constraints. The characteristics of the benchmark set
including required real-time waits and program sizes are shown in
Table~\ref{fig:benchlist}.  {\renewcommand{\arraystretch}{0.9}
%\renewcommand{\texttt}[1] {\begin{lstlisting}[language=Java,morekeywords={delay}] #1 \end{lstlisting}}
\begin{table}[h!]
	\scriptsize
	\centering
	\caption{An overview of the benchmark programs}
	\label{fig:benchlist}
	\scalebox{0.9}{
		\begin{tabular}{llp{0.5cm}p{1cm}}
\toprule
\emph{Benchmark}       & \emph{Delay constructs used in each CD} & \emph{\# of CDs} & 
\emph{Lines of code}        
\\ \midrule
HRTCS                  & \texttt{wait\_inbetween(50.3..200.3 ms)} ($CD1$) & 2                  & 56
\\ \midrule
\multirow{5}{*}{Motor} & \texttt{wait\_exactly(2.4 ms)}        & \multirow{5}{*}{1} & \multirow{5}{*}{116} \\ 
											 &
                                                                                         \texttt{wait\_exactly(1.667
                                                                                           ms)}      &                    &                      \\ 
											 &
                                                                                    \texttt{wait\_exactly(0.05
                                                                                           ms)}       &                    &                      \\ 
											 &
                                                                                    \texttt{wait\_exactly(0.3
                                                                                           ms)}        &                    &                      \\ 
											 &
                                                                                         \texttt{wait\_exactly(0.733
                                                                                           ms)}      &                    &
\\ \midrule
\multirow{3}{*}{Printer}
& \texttt{wait\_inbetween(2.7..10.6 ms)} ($CD1$)      & \multirow{3}{*}{2}        & \multirow{3}{*}{39} \\
& \texttt{wait\_inbetween(1.2..30.3 ms)} ($CD2$)		  &														&											\\
& \texttt{wait\_inbetween(5.6..100.2 ms)} ($CD2$)     &														&											
\\ \midrule
\multirow{2}{*}{AECS} & \texttt{wait\_atleast(10000 ms)} ($CD1$) 
& \multirow{2}{*}{2} & \multirow{2}{*}{267} \\
& \texttt{wait\_exactly(10000 ms)} ($CD2$) & &
\\ \bottomrule
\end{tabular}
}
\end{table}
}
{\renewcommand{\arraystretch}{0.6}
\begin{table*}[t!]
\centering
	\caption{Actual delays obtained for wait constructs in the
benchmark programs based on their BCRT and WCRT}
\begin{tabular}{ c c l l l l l l }
	\toprule
	\multicolumn{2}{c}{} & \emph{HRTCS} & \emph{Motor} & \emph{Printer/CD1} &
	\emph{Printer/CD2} & \emph{AECS/CD1} &	\emph{AECS/CD2} \\ 
	\midrule
	\multicolumn{2}{c}{\emph{BCRT}}	& 0.0.0396 ms	& 0.0425 ms & 0.0190 ms &
	0.0263 ms & 0.1784 ms & 0.1284 ms\\ 
	\cmidrule(r){1-8}
	\multicolumn{2}{c}{\emph{WCRT}} &0.1152 ms & 0.1096 ms & 0.0647 ms
	& 0.0491 ms & 0.3476 ms & 0.2991 ms\\ 
	\cmidrule(r){1-8}

@@

AECS (Access and Environmental Control System) is an application that
controls an intelligent room environment described
in~\cite{aecs_ispa}. The system interacts with two external timers
through input/output signals in order to measure how long the entrance
door to the room is being opened to detect intruders.  The program has
been modified such that the wait constructs are used instead of external
timers as shown in Figure~\ref{delaycode}. In the original AECS, the
system needed to trigger an external timer by emitting the output signal
in order to wait for a specific amount of time (e.g.
\texttt{TimerTriggered} in Figure~\ref{delaycode:a}). When the timer
expires, it notifies the system through the input signal
\texttt{TimeOut}, which is polled using the \texttt{await}
statement. With the new wait constructs, these timers are not necessary
and waits can be expressed directly in the code as shown in
Figure~\ref{delaycode:b}.

### 2014-01-27 introduction.tex
Added proofs, can talk a bit more about completeness and soundness

Reactive languages~\cite{gber931,amal10} are a class of programming
languages used for designing and implementing reactive systems, which
continuously respond to input from their environment. These languages
have been successfully used in programming a plethora of systems such as
fly-by-wire in Airbus~\cite{eairbus}, security surveillance
systems~\cite{amal121}, etc. Usually these systems also need to meet
real-time constraints imposed by the environment. Yet, these languages
do not support describing real-time statements as first class language
constructs.  For example, one cannot describe a simple real-time delay
(postponement) of an operation for 0.2 ms. These languages are based on
formal semantics, essential for the formal reasoning about and
verification of correctness of functional properties of the developed
programs, but leave nonfunctional properties such as timing behavior as
an implementation detail~\cite{boldt07}. Arguably, rightly so, because
physical time cannot be incorporated without information about the
underlying execution platform.  But, one can still reason in terms of
logical time. These languages support discrete logical clock rather than
a discrete physical (real-time) clock. The real-time period of the
discrete logical clock (also referred to as a logical tick) is not fixed
and it is determined by the responsiveness of the program to the input
signals. Unlike a discrete physical clock, which has fixed real-time
period, the period of the logical clock is elastic. Although the timing
model with logical clock has worked well for reactive synchronous/GALS
languages in designing discrete systems, there is a need to introduce
real-time since, often, implementation models are in the continuous or
discrete real-time domain~\cite{DBLP:journals/pieee/SifakisTY03}.
Moreover, adding timing capabilities as first class language constructs
create additional opportunity for formally verifying functional and
non-functional requirements before system deployment.

@@

The major motivations for introducing these real-time \textit{wait}
mechanisms and at the same time \textbf{contributions} of the paper are:
(1) these waiting mechanisms allow real-time based synchronization
between concurrent behaviors, (2) the waiting model relies on the use of
relative instead of absolute real-time, i.e., a wait in selected time
units is counted from the currently executing SystemJ statement, (3)
these waiting mechanisms allow mixing behaviors with real-time features
and others with only logical time and (4) real-time wait does not affect
the model of logical time in SystemJ program as long as the amount of
delay is within boundaries that can be determined statically by program
analysis. All these as well as semantics of the delay construct are
discussed and illustrated in the rest of the paper. Our contributions
can be further refined as follows:
\begin{enumerate*}
\item Real-time is modeled in the $\mathbb{Q}^{>0}$ domain of numbers,
 i.e., non-negative rational numbers.
\item We consider the property of \textit{non
    maximal-parallelism}. \textit{Non-maximal parallelism:} There are
  not always sufficient resources for processes to execute, so
  effectively processes are allocated and scheduled with different
  optimization criteria -- reducing computation time, reducing power
  consumption, etc. This allocation and scheduling is an integral part
  of any language compilation and needs to be considered when
  introducing real-time delays.
\item We are the first, to our knowledge, to allow specification of
  both; exact and non-exact real-time in GALS and synchronous programs.
\item We elucidate upon a key insight relating real-time and
  logical-time -- \textit{the resolution of any real-time statement is
    dependent on the worst case logical tick time} -- an aspect that
  seems to have escaped in the current literature of real-time in
  synchronous/GALS languages.
\end{enumerate*}

The rest of the paper is arranged as follows:
Section~\ref{sec:background} gives the background on SystemJ and related
techniques required for the rest of the
paper. Section~\ref{sec:intr-real-time} introduces real-time mechanisms
in the GALS paradigm and describes their compilation into logical
time. Section~\ref{sec:experimental-results} gives the experimental
results for a set of real-time
applications. Section~\ref{sec:related-work} positions this work in
relation to other approaches to introducing real-time. Finally, we
conclude in Section~\ref{sec:concl-future-work}.

>>>>

Reactive languages~\cite{gber931,amal10} are a class of programming
languages used for designing and implementing reactive systems, which
continuously respond to input from their environment. These languages
have been successfully used in programming a plethora of systems such as
fly-by-wire in Airbus~\cite{eairbus}, security surveillance
systems~\cite{amal121}, etc. Usually these systems also need to meet
real-time constraints imposed by the environment. Yet, these languages
do not support describing real-time statements as first class language
constructs.  For example, one cannot describe a simple real-time delay
(postponement) of an operation for 0.2 ms. These languages are based on
formal semantics, essential for the formal reasoning about and
verification of correctness of functional properties of the developed
programs, but leave nonfunctional properties such as timing behavior as
an implementation detail~\cite{boldt07}. Arguably, rightly so, because
physical time cannot be incorporated without information about the
underlying execution platform.  But, one can still reason in terms of
logical time. These languages support discrete logical clock rather than
a discrete physical (real-time) clock. The real-time period of the
discrete logical clock (also referred to as a logical tick) is not fixed
and it is determined by the responsiveness of the program to the input
signals. Unlike a discrete physical clock, which has fixed real-time
period, the period of the logical clock is elastic. Although the timing
model with logical clock has worked well for reactive synchronous/GALS
languages in designing discrete systems, there is a need to introduce
real-time, since often, implementation models are in the continuous or
discrete real-time domain~\cite{DBLP:journals/pieee/SifakisTY03}.
Moreover, adding timing capabilities as first class language constructs
create additional opportunity for formally verifying functional and
non-functional requirements before system deployment.

@@

The major motivations for introducing these real-time \textit{wait}
mechanisms and at the same time \textbf{contributions} of the paper are:
(1) these waiting mechanisms allow real-time based synchronization
between concurrent behaviors, (2) the waiting model relies on the use of
relative instead of absolute real-time, i.e., a wait in selected time
units is counted from the currently executing SystemJ statement, (3)
these waiting mechanisms allow mixing behaviors with real-time features
and others with only logical time and (4) real-time wait does not affect
the model of logical time in SystemJ program as long as the amount of
delay is within boundaries that can be determined statically by program
analysis. All these as well as semantics of the wait constructs are
discussed and illustrated in the rest of the paper. Our contributions
can be further refined as follows:
\begin{enumerate*}
\item Real-time is modeled in the $\mathbb{Q}^{>0}$ domain of numbers,
 i.e., non-negative rational numbers.
\item We consider the property of \textit{non
    maximal-parallelism}. \textit{Non-maximal parallelism:} There are
  not always sufficient resources for processes to execute, so
  effectively processes are allocated and scheduled with different
  optimization criteria -- reducing computation time, reducing power
  consumption, etc. This allocation and scheduling is an integral part
  of any language compilation and needs to be considered when
  introducing real-time waits.
\item We are the first, to our knowledge, to allow specification of
  both; exact and non-exact real-time in GALS and synchronous programs.
\item We elucidate upon a key insight relating real-time and
  logical-time -- \textit{the resolution of any real-time statement is
    dependent on the worst case logical tick time} -- an aspect that
  seems to have escaped in the current literature on incorporation of
  real-time in synchronous/GALS languages.
\end{enumerate*}

The rest of the paper is arranged as follows:
Section~\ref{sec:background} gives the background on SystemJ and related
techniques required for the rest of the
paper. Section~\ref{sec:intr-real-time} introduces real-time mechanisms
in the GALS paradigm and describes their compilation into logical
time. Section~\ref{sec:proof-correctness} sketches the correctness
proofs for the introduced
algorithms. Section~\ref{sec:experimental-results} gives the
experimental results for a set of real-time applications. In
Section~\ref{sec:disc-perc-limit} we discuss the advantages and
limitations of the proposed approach. Section~\ref{sec:related-work}
positions this work in relation to other approaches to introducing
real-time. Finally, we conclude in Section~\ref{sec:concl-future-work}.
